---
title: "The Good, the Bad and the Ugly: Daten effektiv visualisieren und kommunizieren"
author: "Dr. Shirin Elsinghorst"
bibliography: references.bib
date: "8/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align="center")
```

```{r libraries}
library(tidyverse)

library(ggExtra)
library(ragg)
library(ggalluvial)
library(treemapify)
library(ggalt)

library(palmerpenguins)
```

## Dataset

```{r}
head(penguins)
#head(penguins_raw)
```

```{r eval=FALSE, echo=FALSE}
p <- penguins %>%
  ggplot(aes(x = bill_length_mm, y = flipper_length_mm,
              color = species)) +
    geom_point()

pngfile <- fs::path(knitr::fig_path(),  "scaling.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 300, scaling = 3)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

## Colors

- set [colorblind-friendly palettes](https://jfly.uni-koeln.de/color/)

```{r}
# The palette with grey:
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# The palette with black:
cbp2 <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

library(plotrix)
sliceValues <- rep(10, 8) # each slice value=10 for proportionate slices
(
  p <- pie3D(sliceValues, 
      explode=0, 
      theta = 1.2, 
      col = cbp1, 
      labels = cbp1, 
      labelcex = 0.9,
      shade = 0.6,
      main = "Colorblind\nfriendly palette")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "colors.png")
agg_png(pngfile, width = 40, height = 40, units = "cm", res = 100, scaling = 4)
pie3D(sliceValues, 
      explode=0, 
      theta = 1.2, 
      col = cbp1, 
      labels = cbp1, 
      labelcex = 0.9,
      shade = 0.6,
      main = "Colorblind\nfriendly palette")
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
ggplot <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = cbp1) +
  scale_fill_manual(values = cbp1) +
  theme_bw()
```

## Main Diagram types

https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf

### Point chart

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = bill_length_mm, y = flipper_length_mm)) +
    geom_jitter(alpha = 0.5) +
    facet_wrap(vars(species), ncol = 3) +
    scale_x_reverse() +
    scale_y_reverse() +
    labs(x = "Bill length (mm)", 
         y = "Flipper length (mm)",
         size = "body mass (g)",
        title = "Scatterplot", 
        subtitle = "Penguins bill v. flipper length by species",
        caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "gg_demo_1d.png")
agg_png(pngfile, width = 60, height = 30, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = bill_length_mm, y = flipper_length_mm,
              color = species, shape = species)) +
    geom_point(alpha = 0.7) +
    labs(x = "Bill length (mm)", 
         y = "Flipper length (mm)",
        title = "Scatterplot", 
        subtitle = "Penguins bill v. flipper length by species",
        caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "scatterplot.png")
agg_png(pngfile, width = 60, height = 36, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

- Jitter with smoothing line

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = bill_length_mm, y = flipper_length_mm,
              color = species, shape = species)) +
    geom_jitter(alpha = 0.5) +
    geom_smooth(method = "loess", se = TRUE) +
    facet_wrap(vars(species), nrow = 3) +
    labs(x = "Bill length (mm)", 
         y = "Flipper length (mm)",
        title = "Scatterplot with smoothing line", 
        subtitle = "Penguins bill v. flipper length by species with loess smoothing line",
        caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "scattersmooth.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
p <- penguins %>%
  remove_missing() %>%
  filter(species == "Adelie") %>%
  ggplot(aes(x = bill_length_mm, y = flipper_length_mm)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = TRUE) +
    labs(x = "Bill length (mm)", 
         y = "Flipper length (mm)",
        title = "Scatterplot with smoothing line", 
        subtitle = "Penguins bill v. flipper length by species with\nloess smoothing line, histogram & density distribution",
        caption = "Source: https://github.com/allisonhorst/palmerpenguins")

#(p <- ggMarginal(p, type = "densigram", fill = "transparent"))
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "scatterhist.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
ggMarginal(p, type = "densigram", fill = "transparent")
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Bubble chart

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = bill_length_mm, y = flipper_length_mm,
              color = species, shape = species, size = body_mass_g)) +
    geom_point(alpha = 0.5) +
    labs(x = "Bill length (mm)", 
         y = "Flipper length (mm)",
        title = "Bubble plot", 
        size = "body mass (g)",
        subtitle = "Penguins bill v. flipper length by species;\nsize indicates body mass in grams",
        caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "bubble.png")
agg_png(pngfile, width = 60, height = 40, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Line chart

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    filter(species == "Adelie") %>%
    ggplot(aes(x = bill_length_mm, y = flipper_length_mm,
               color = sex)) +
    geom_line() +
    geom_point() +
    labs(x = "Bill length (mm)", 
         y = "Flipper length (mm)",
        title = "Line plot", 
        subtitle = "Adelie penguins bill v. flipper length by sex",
        caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "linechart.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Correlation plots / heatmap

```{r}
mat <- penguins %>%
  remove_missing() %>%
  select(bill_depth_mm, bill_length_mm, body_mass_g, flipper_length_mm)

cormat <- round(cor(mat), 2)
cormat[upper.tri(cormat)] <- NA

cormat <- cormat %>%
  as_data_frame() %>%
  mutate(x = colnames(mat)) %>%
  gather(key = "y", value = "value", bill_depth_mm:flipper_length_mm)

(
  p <- cormat %>%
    remove_missing() %>%
    arrange(x, y) %>%
    ggplot(aes(x = x, y = y, fill = value)) + 
    geom_tile() +
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
     midpoint = 0, limit = c(-1,1), space = "Lab", 
     name = "Pearson\nCorrelation") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    coord_fixed() +
      labs(x = "", 
           y = "",
          title = "Correlation heatmap", 
          subtitle = "Correlation btw. penguins' traits",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "corplot.png")
agg_png(pngfile, width = 60, height = 60, units = "cm", res = 100, scaling = 5)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Bar chart

- per default: counts

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = species,
               fill = sex)) +
    geom_bar() +
      labs(x = "Species", 
           y = "Counts",
          title = "Barchart", 
          subtitle = "Counts of male & female penguins per species in study",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "barchart.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = species,
               fill = sex)) +
    geom_bar(position = 'dodge') +
      labs(x = "Species", 
           y = "Counts",
          title = "Barchart", 
          subtitle = "Counts of male & female penguins per species in study",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "barchart_dodge.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

- alternative: set y-values

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    group_by(species, sex) %>%
    summarise(mean_bmg = mean(body_mass_g),
              sd_bmg = sd(body_mass_g)) %>%
    ggplot(aes(x = species, y = mean_bmg,
               fill = sex)) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_errorbar(aes(ymin = mean_bmg - sd_bmg, 
                      ymax = mean_bmg + sd_bmg), 
                  width = 0.2,
                 position = position_dodge(0.9)) +
      labs(x = "Species", 
           y = "Mean body mass (in g)",
          title = "Barchart", 
          subtitle = "Mean body mass of male & female penguins per species\nwith standard deviation",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "barchart_dodge_bmg.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Boxplot

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = species, y = body_mass_g,
               fill = sex)) +
    geom_boxplot() +
      labs(x = "Species", 
           y = "Body mass (in g)",
          title = "Boxplot", 
          subtitle = "Body mass of three penguin species per sex",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "boxplot.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

- with points

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = species, y = body_mass_g,
               fill = sex, color = sex)) +
    geom_boxplot(alpha = 0.5, notch = TRUE) +
    geom_jitter(alpha = 0.5, position=position_jitter(0.3)) +
      labs(x = "Species", 
           y = "Body mass (in g)",
          title = "Boxplot with points (dotplot)", 
          subtitle = "Body mass of three penguin species per sex",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "dotplot.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Violin plot

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = species, y = body_mass_g,
               fill = sex)) +
    geom_violin(scale = "area") +
      labs(x = "Species", 
           y = "Body mass (in g)",
          title = "Violinplot", 
          subtitle = "Body mass of three penguin species per sex",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "violinplot.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    ggplot(aes(x = species, y = body_mass_g,
               fill = sex, color = sex)) +
    geom_dotplot(method = "dotdensity", alpha = 0.7,
                 binaxis = 'y', stackdir = 'center',
                 position = position_dodge(1)) +
      labs(x = "Species", 
           y = "Body mass (in g)",
          title = "Violinplot with points (dotplot)", 
          subtitle = "Body mass of three penguin species per sex",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "violindotplot.png")
agg_png(pngfile, width = 60, height = 40, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Pie chart

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    group_by(species, sex) %>%
    summarise(n = n()) %>%
    mutate(freq = n / sum(n),
           percentage = freq * 100) %>%
    ggplot(aes(x = "", y = percentage,
               fill = sex)) +
    facet_wrap(vars(species), nrow = 1) +
    geom_bar(stat = "identity", alpha = 0.8) +
    coord_polar("y", start = 0) +
      labs(x = "", 
           y = "Percentage",
          title = "Piechart", 
          subtitle = "Percentage of male v. female penguins per species in study",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "piechart.png")
agg_png(pngfile, width = 60, height = 40, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Alluvial charts

```{r}
(
  p <- as.data.frame(UCBAdmissions) %>%
    ggplot(aes(y = Freq, axis1 = Gender, axis2 = Dept)) +
    geom_alluvium(aes(fill = Admit), width = 1/12) +
    geom_stratum(width = 1/12, fill = "black", color = "grey") +
    geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
    scale_x_discrete(limits = c("Gender", "Dept"), expand = c(.05, .05)) +
      labs(x = "", 
           y = "Frequency",
          title = "Alluvial chart", 
          subtitle = "UC Berkeley admissions and rejections, by sex and department",
          caption = "Source: Bickel et al. (1975)\nSex bias in graduate admissions: Data from Berkeley. Science, 187, 398–403.")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "alluvial.png")
agg_png(pngfile, width = 60, height = 40, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Maps

https://shiring.shinyapps.io/wgs_app

### Treemaps

```{r}
(
  p <- as.data.frame(UCBAdmissions) %>%
    group_by(Admit, Gender) %>%
    summarise(sum_freq = sum(Freq)) %>%
    ggplot(aes(area = sum_freq, fill = sum_freq, label = Gender, 
               subgroup = Admit)) +
    geom_treemap() +
    geom_treemap_subgroup_border() +
    geom_treemap_subgroup_text(place = "centre", grow = T, alpha = 0.5, colour =
                             "black", fontface = "italic", min.size = 0) +
    geom_treemap_text(colour = "white", place = "centre", reflow = T) +
    scale_fill_gradient2(low = "#999999", high = "#E69F00", mid = "white", midpoint = 1000, space = "Lab", 
     name = "Sum of\nfrequencies") +
      labs(x = "", 
           y = "",
          title = "Treemap", 
          subtitle = "UC Berkeley admissions and rejections by sex",
          caption = "Source: Bickel et al. (1975)\nSex bias in graduate admissions: Data from Berkeley. Science, 187, 398–403.")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "treemap.png")
agg_png(pngfile, width = 60, height = 40, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Dumbbell plots

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    group_by(year, species, sex) %>%
    summarise(mean_bmg = mean(body_mass_g)) %>%
    mutate(species_sex = paste(species, sex, sep = "_"),
         year = paste0("year_", year)) %>%
    spread(year, mean_bmg) %>%
    ggplot(aes(x = year_2007, xend = year_2009, 
               y = reorder(species_sex, year_2009))) +
    geom_dumbbell(color = "#999999", 
                      size_x = 3, 
                      size_xend = 3,
                      #Note: there is no US:'color' for UK:'colour' 
                      # in geom_dumbbel unlike standard geoms in ggplot()
                      colour_x = "#999999",
                      colour_xend = "#E69F00") +
      labs(x = "Body mass (g)", 
           y = "Species & sex",
          title = "Dumbbell plot", 
          subtitle = "Penguin's change in body mass from 2007 to 2009",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "dumbbell.png")
agg_png(pngfile, width = 60, height = 30, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Slope charts

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    group_by(year, species, sex) %>%
    summarise(mean_bmg = mean(body_mass_g)) %>%
    ggplot(aes(x = year, y = mean_bmg, group = sex,
               color = sex)) +
    facet_wrap(vars(species), nrow = 3) +
    geom_line(alpha = 0.6, size = 2) +
    geom_point(alpha = 1, size = 3) +
    scale_x_continuous(breaks=c(2007, 2008, 2009)) +
      labs(x = "Year", 
           y = "Body mass (g)",
           color = "Sex",
          title = "Slope chart", 
          subtitle = "Penguin's change in body mass from 2007 to 2009",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "slope.png")
agg_png(pngfile, width = 50, height = 60, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Stacked area charts

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    group_by(year, species, sex) %>%
    summarise(mean_bmg = mean(body_mass_g)) %>%
    ggplot(aes(x = year, y = mean_bmg, fill = sex)) +
    facet_wrap(vars(species), nrow = 3) +
    geom_area(alpha = 0.6, size=.5, color = "white") +
    scale_x_continuous(breaks=c(2007, 2008, 2009)) +
      labs(x = "Year", 
           y = "Mean body mass (g)",
           color = "Sex",
          title = "Stacked area chart", 
          subtitle = "Penguin's change in body mass from 2007 to 2009",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "area.png")
agg_png(pngfile, width = 50, height = 60, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Lolliplot chart

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    group_by(year, species, sex) %>%
    summarise(mean_bmg = mean(body_mass_g)) %>%
    mutate(species_sex = paste(species, sex, sep = "_"),
         year = paste0("year_", year)) %>%
    spread(year, mean_bmg) %>%
    ggplot() +
    geom_segment(aes(x = reorder(species_sex, -year_2009), xend = reorder(species_sex, -year_2009), 
                   y = 0, yend = year_2009),
                 color = "#999999", size = 1) +
    geom_point(aes(x = reorder(species_sex, -year_2009), y = year_2009),
               size = 4, color = "#E69F00") +
    coord_flip() +
      labs(x = "Species & sex", 
           y = "Body mass (g)",
          title = "Lollipop chart", 
          subtitle = "Penguin's body mass in 2009",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "lollipop.png")
agg_png(pngfile, width = 60, height = 30, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Dendrograms

```{r}
library(ggdendro)
library(dendextend)
```

```{r}
penguins_hist <- penguins %>%
  filter(sex == "male") %>%
  select(species, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>%
  group_by(species) %>% 
  sample_n(10) %>%
  as.data.frame()
rownames(penguins_hist) <- paste(penguins_hist$species, seq_len(nrow(penguins_hist)), sep = "_")

penguins_hist <- penguins_hist %>%
  select(-species) %>%
  remove_missing()
```

```{r}
#hc <- hclust(dist(penguins_hist, method = "euclidean"), method = "ward.D2")
#ggdendrogram(hc)

# Create a dendrogram and plot it
dend <- penguins_hist %>%  
  scale %>% 
  dist(method = "euclidean") %>%
  hclust(method = "ward.D2") %>% 
  as.dendrogram
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "dendrogram.png")
agg_png(pngfile, width = 60, height = 35, units = "cm", res = 100, scaling = 4)
dend %>% 
  set("labels_col", value = cbp1, k=3) %>% 
  plot(main = "Dendrogram")
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Waterfall charts

```{r}
library(waterfall)
```

```{r}
(
  p<- jaquith %>%
    arrange(score) %>%
    add_row(factor = "Total", score = sum(jaquith$score)) %>%
    mutate(factor = factor(factor, levels = factor),
                           id = seq_along(score)) %>%
    mutate(end = cumsum(score),
           start = c(0, end[-length(end)]),
           start = c(start[-length(start)], 0),
           end = c(end[-length(end)], score[length(score)]),
           gr_col = ifelse(factor == "Total", "Total", "Part")) %>%
    ggplot(aes(x = factor, fill = gr_col)) + 
      geom_rect(aes(x = factor,
                    xmin = id - 0.45, xmax = id + 0.45, 
                    ymin = end, ymax = start)) +
      theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1),
            legend.position = "none") +
        labs(x = "", 
             y = "Amount",
            title = "Waterfall chart", 
            subtitle = "Sample business-adjusted risk from Security Metrics",
            caption = "Andrew Jaquith, Security Metrics: Replacing Fear, Uncertainty, and Doubt\n(Boston: Addison-Wesley Professional, 2007), 170-171.")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "waterfall.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Biplots

```{r}
library(ggfortify)
```

```{r}
penguins_prep <- penguins %>%
  remove_missing() %>%
  select(bill_length_mm:body_mass_g)

penguins_pca <- penguins_prep %>%
  prcomp(scale. = TRUE)
```

```{r}
penguins_km <- penguins_prep %>%
  kmeans(3)
```

```{r}
(
  p <- autoplot(penguins_pca, 
                data = penguins %>% remove_missing(), 
                colour = 'species',
                shape = 'species',
                loadings = TRUE, 
                loadings.colour = 'blue',
                loadings.label = TRUE, 
                loadings.label.size = 3) +
      scale_color_manual(values = cbp1) +
  scale_fill_manual(values = cbp1) +
  theme_bw() +
            labs(
            title = "Biplot PCA", 
            caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "biplot_pca.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
(
  p <- autoplot(penguins_km, 
                data = penguins %>% remove_missing(), 
                colour = 'species',
                shape = 'species',
                frame = TRUE, frame.type = 'norm') +
      scale_color_manual(values = cbp1) +
  scale_fill_manual(values = cbp1) +
  theme_bw() +
            labs(
            title = "Biplot k-Means clustering", 
            caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "biplot_km.png")
agg_png(pngfile, width = 60, height = 50, units = "cm", res = 100, scaling = 4)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

### Radar charts, aka star chart, aka spider plot

https://www.data-to-viz.com/caveat/spider.html

```{r}
library(ggiraphExtra)
```

```{r}
(
  p <- penguins %>%
    remove_missing() %>%
    select(-island, -year) %>%
    ggRadar(aes(x = c(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g), 
                group = species,
                colour = sex, facet = sex), 
            rescale = TRUE, 
            size = 1, interactive = FALSE, 
            use.label = TRUE) +
     scale_color_manual(values = cbp1) +
  scale_fill_manual(values = cbp1) +
  theme_bw() +
     scale_y_discrete(breaks = NULL) + # don't show ticks
      labs(
          title = "Radar/spider/star chart", 
          subtitle = "Body mass of male & female penguins per species",
          caption = "Source: https://github.com/allisonhorst/palmerpenguins")
)
```

```{r eval=FALSE, echo=FALSE}
pngfile <- fs::path("Figs",  "radar.png")
agg_png(pngfile, width = 60, height = 40, units = "cm", res = 100, scaling = 3)
plot(p)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

---

```{r}
devtools::session_info()
```

